How to improve your relationship with your future self.

### 2016-06-24 workflow.Rmd
some more

---
title: How to improve your relationship with your future self.
author:
- name: Jake Bowers
  affiliation: <jwbowers@illinois.edu>,  University of Illinois at Urbana-Champaign, Departments of Political Science and Statistics; Fellow, White House Social and Behavioral Sciences Team. Nothing in this paper reflect the official position of the US Government. Many thanks to Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn and Cara Wong for direct help on this document and to Mika LaVaque-Manty and Ben Hansen for many useful discussions on this topic. The source code for this document may be freely downloaded and modified from <https://github.com/jwbowers/workflow>.
- name: Maarten Voors
  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group. Special thanks to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during 2016.
bibliography: references.bib
csl: american-political-science-review.csl
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
    pdf_document:
      keep_tex: true
      number_sections: true
      fig_width: 7
      fig_height: 7
      fig_caption: true
      template: bowersarticletemplate.latex
      pandoc_args: [
      "-M", "secPrefix=\\S",
      "--filter", "/usr/local/bin/pandoc-crossref"
      ]
      md_extensions: +autolink_bare_uris+ascii_identifiers+tex_math_single_backslash
...

@@

How long does it take from planning a study to publication? And subsequently, how much time passes to the first reproduction of a study? Is three years too short? Is ten years too long? We suspect that few of our colleagues in the social and behavioral sciences conceive of, field, write, and publish a data driven study faster than about three years. We also suspect that, if some other scholar decides to re-examine the analyses of a published study, it will occur after publication. Moreover, this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis. Other times, this process is organized more broadly (cite IIIE program, cite Open Science Foundation on psychology, cite the Camerer 2016 Science paper?, cite the cool "Crowdsourcing Analysis" project by Silberzahn and Uhlmann Nature 2015?).]

@@

@@

Finally, coding helps us to avoid making mistakes. For example, say in our example above we are interested in how many people protest. We may use a data file containing all protests for several years. We can then create a table with the number of protesters by copying the results manually to a document you use to write your paper. In copying we can make mistakes. So a better approach is to ask R to automatically create the table (in the format you like, with horizontal lines, 3 decimals, a title, etc) and output this in whatever file type you need (.pdf, .rtf, etc). Now when we obtain new data (for additional years, or updated figures for existing data) a new table can be created quickly, so we make less mistakes and save time!

[NOTE we could include some code here to provide an example of automation of results?]

@@

 An excellent, simple, and robust version
control system is to rename your files with the date and time of
saving them: yyyymmdd_project.docx (for changes within the same day, just before a deadline for example, you may even add a time, so it becomes 20160623_4PM_inequalitypaper.docx).  Be sure to include year in the file names --- remember, the life of an idea is
measured in years. If you are wise enough to have saved your documents
as plain text then you can easily compare documents using the many
utilities available for comparing text files.^[Adobe Acrobat
  allows one to compare differences in pdf files. OpenOffice supports
  a 'Compare Documents 'option. Word now does the same. And Google Docs will report on the
  version history of a document.] When you reach certain milestones
you can rename the file accordingly: thedocAPSA2009.tex --- for the
one sent to discussants at APSA --- or thedocAPSR2015.tex --- for the
version eventually sent to the APSR six years after you presented it
at APSA. The formal version control systems mentioned above all
allow this kind of thing and are much more elegant and capable, but
you can do it by hand too as long as you don't mind taking up a lot of
disk space and having many 'thedoc... ' files around. Indeed, the number of files can add up quickly (especially around submission time!) and you may want to create an "0_Archive" folder to store older versions. If you do
version control by hand, spend a little extra time to ensure that you
do not clobber files when you make mistakes typing in the
file-names. And, if you find yourself spending extra time reconciling
changes made by different collaborators by hand, remember this is a
task that modern version control systems take care of quickly and
easily.

@@

Anyone writing custom code should worry about getting it right. The more code one writes, the more time one has to appreciate problems arising from bugs, errors, and typos in data analysis and code. Mistakes will always be made we should just recognize we need more standardization of practice to minimise the mistakes and help others find them. One thing we can do is to include testing in our coding.

@@

Ack! we mistyped '+' for '*'. Good thing I wrote the
test!^[A more common example of this kind of testing occur
  everyday when we recode variables into new forms but look at a
  crosstab of the old vs. new variable before proceeding.]

@@

# Copy and improve on others' examples.

[IS THIS SECTION MORE MORE ON CREATING AN INTEGRATED REPRODUCIBLE WORKFLOW RATHER THAN COPYING AND IMPROVING?]
Lots of people are thinking about 'reproducible research', creating a 'reproducible workflow' and
'literate programming' these days. Google those terms. Of course, the
devil is in the details: Here I list a few of my own attempts at enabling
reproducible research. You'll find many other inspiring examples on
the web. Luckily, the open source ethos aligns nicely with academic
incentives, so we are beginning to find more and more people offering
their files online for copying and improvement. By the way, if you do
copy and improve, it is polite to alert the person from whom you made
the copy about your work.

We have experimented with several systems so far: (1) we wrote this paper in R markdown (Jake) and Atom (Maarten) and organized our files and collaboration on Github. The source code can be accessed, downloaded and modified from <https://github.com/jwbowers/workflow>. The program also allowed us to send notes to each other. The nice thing about github is that is enables you to make "releases" of a project which enable you to smooth the reproduction of all of the
products of your research (simulations, data analyses, tables, figures, etc..). Similar features are offered by the Open Science Framework (https://osf.io/), Maarten recently participated in a workshop on research transparency organized by the Berkeley Institute for Transparency in the Social Sciences (www.bitss.org), all course material was accessible through OSF (https://osf.io/qh2nr/). Both platforms, OSF and Github nicely integrate and a key benefit is that they have been created to remain for a long time (unlike perhaps your university drive) JAKES PROJECT; (2) for one paper Jake simply included a Sweave document and data files into a compressed archive [@bowers2005dataverse] [SAY ABIT MORE ABOUT WHAT SWEAVE IS IN PRINCIPLE]; (3) for another more computing intensive paper Jake
assembled a set of files that enabled reproduction of our results using the
`make` system [@bowers2008dataverse] [EXPLAIN A BIT MORE OR INCLUDE REF TO WEBSITE?]; (4) Jake also tried the "compendium"
approach [@gentleman2005reproducible,@gentleman2007statistical] which embeds an
academic paper within the R package system [@bowers2011dataverse]; The benefit of the compendium approach is that one is not required to have
access to a command line for `make`: the compendium is downloadable from within
R using `install.packages()` and is viewable using the `vignette()` function in
any operating system than runs R.^[Notice that Jake's reproduction archives and/or
instructions for using them are hosted on the
[Dataverse](http://thedata.org/book/learn-about-project), which is another
system designed to enhance academic collaboration across time and space.]  The
idea that one ought to be able to install and run and use an academic paper
just as one installs and uses statistical software packages is very attractive,
and we anticipate that it will become ever easier to turn papers into R packages
as creative and energetic folks turn their attention to the question of
reproducible research.^[See for example <http://r-pkgs.had.co.nz/>]
Finally, (5) Maarten has been using Dropbox for most of his collaborative projects (https://www.dropbox.com/). ALREADY ABOVE? NOTE BACK

There are some new noteworthy developments also. For example, and similar to R Markdown/KnitR in R Studio, there is Markdoc for STATA users (https://github.com/haghish/MarkDoc). People at UC Berkeley have been developing a is a web app, Jupyter Notebook (http://jupyter.org/), that enables you to make and share documents that include text, code, equations and visualizations in one INCLUDE REF TO GITHUB - USING PYTHON. For researchers working across difference platforms, there is Docker (https://www.docker.com/).

@@

 As we copy and
improve upon each other's modes of doing work we enhance our collective
ability to believe each other and for future scholars to believe us, too.
[I FEEL LIKE THE COPY COMMENT IS PERHAPS ANOTHER POINT TO MAKE (OR LEAVE)]

>>>>

---
title: How to improve your relationship with your future self.
author:
- name: Jake Bowers
  affiliation: <jwbowers@illinois.edu>,  University of Illinois at Urbana-Champaign, Departments of Political Science and Statistics; Fellow, White House Social and Behavioral Sciences Team. Nothing in this paper reflect the official position of the US Government. Many thanks to Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn and Cara Wong for direct help on this document and to Mika LaVaque-Manty and Ben Hansen for many useful discussions on this topic. The source code for this document may be freely downloaded and modified from <https://github.com/jwbowers/workflow>.
- name: Maarten Voors
  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group. Special thanks, the EGAP Learning Days 2016 participants in Santiago Chile,  to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
  <!-- Separate the thanks sections from affiliation and integrate? Wasnt sure how to do this here-->
bibliography: references.bib
csl: american-political-science-review.csl
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
    pdf_document:
      keep_tex: true
      number_sections: true
      fig_width: 7
      fig_height: 7
      fig_caption: true
      template: bowersarticletemplate.latex
      pandoc_args: [
      "-M", "secPrefix=\\S",
      "--filter", "/usr/local/bin/pandoc-crossref"
      ]
      md_extensions: +autolink_bare_uris+ascii_identifiers+tex_math_single_backslash
...

@@

How long does it take from planning a study to publication? And subsequently, how much time passes to the first reproduction of a study? Is three years too short? Is ten years too long? We suspect that few of our colleagues in the social and behavioral sciences conceive of, field, write, and publish a data driven study faster than about three years. We also suspect that, if some other scholar decides to re-examine the analyses of a published study, it will occur after publication. Moreover, this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis. Other times, this process is organized more broadly (cite IIIE program, researchers have started to implement psychological experiments using the same protocols across many locations, see Open Science Foundation 2014, 2015, within experimental economics a team did the same, see Camerer et al 2016 Science paper. Another example is to share analysis across many researchers. in a their 'crowdsourcing analysis" Silberzahn and Uhlmann (Nature 2015) asked 29 research teams to run the analysis for a project that looked at the impact of skin color on the probability of getting a foul card in a football match.<!-- Can you do the correct citing here -->]

@@

<!-- JAKE: we dont say anything about pre registration, should we? -->

@@

Finally, coding helps us to avoid making mistakes. For example, in our example above we may be interested in how many people protest. We may use a data file containing all protests for several years. We can then create a table with the number of protesters by copying the results manually to a document you use to write your paper. In copying we can make mistakes. So a better approach is to ask R to automatically create the table (in the format you like, with horizontal lines, 3 decimals, a title, etc) and output this in whatever file type you need (.pdf, .rtf, etc). Now when we obtain new data (for additional years, or updated figures for existing data) a new table can be created quickly, so we make less mistakes and save time!

<!-- Can you include some code here to provide an example of automation of results? -->

@@

An excellent, simple, and robust version control system is to rename your files with the date and time of
saving them: yyyymmdd_project.docx (for changes within the same day, just before a deadline for example, you may even add a time, so it becomes 20160623_4PM_inequalitypaper.docx). This communicates very clear what version you are working. Remember the days when you received a file called inequalityMV4_APSRsubmit_reallyfinal2.tex? That should quickly become something of the past! We find ourselves preaching the gospel to our collaborators frequently and will keep renaming files until a collaborator has been successfully pacified. Be sure to include year in the file names --- remember, the life of an idea is
measured in years. If you are wise enough to have saved your documents
as plain text then you can easily compare documents using the many
utilities available for comparing text files.^[Adobe Acrobat
  allows one to compare differences in pdf files. OpenOffice supports
  a 'Compare Documents 'option. Word now does the same. And Google Docs will report on the
  version history of a document.] When you reach certain milestones
you can rename the file accordingly: thedocAPSA2009.tex --- for the
one sent to discussants at APSA --- or thedocAPSR2015.tex --- for the
version eventually sent to the APSR six years after you presented it
at APSA. The formal version control systems mentioned above all
allow this kind of thing and are much more elegant and capable, but
you can do it by hand too as long as you don't mind taking up a lot of
disk space and having many 'thedoc... ' files around. Indeed, the number of files can add up quickly (especially around submission time!) and you may want to create an "0_Archive" folder to store older versions. If you do
version control by hand, spend a little extra time to ensure that you
do not clobber files when you make mistakes typing in the
file-names. And, if you find yourself spending extra time reconciling
changes made by different collaborators by hand, remember this is a
task that modern version control systems take care of quickly and
easily.

@@

Anyone writing custom code should worry about getting it right. The more code one writes, the more time one has to appreciate problems arising from bugs, errors, and typos in data analysis and code. Mistakes will always be made we should just recognize we need more standardization of practice to minimize the mistakes and help others find them. One thing we can do is to include testing in our coding.

@@

Ack! we mistyped '+' for '*'. Good thing we wrote the
test!^[A more common example of this kind of testing occur
  everyday when we recode variables into new forms but look at a
  crosstab of the old vs. new variable before proceeding.]

@@

# Creating a reproducible workflow

Lots of people are thinking about 'reproducible research', creating a 'reproducible workflow' and
'literate programming' these days. Google those terms. Of course, the
devil is in the details: Here I list a few of my own attempts at enabling
reproducible research. You'll find many other inspiring examples on
the web. Luckily, the open source ethos aligns nicely with academic
incentives, so we are beginning to find more and more people offering
their files online for copying and improvement. By the way, if you do
copy and improve, it is polite to alert the person from whom you made
the copy about your work.

We have experimented with several systems so far: (1) we wrote this paper in R markdown (Jake) and Atom (Maarten) and organized our files and collaboration on Github. The source code can be accessed, downloaded and modified from <https://github.com/jwbowers/workflow>. The program also allowed us to send notes to each other. The nice thing about github is that is enables you to make "releases" of a project which enable you to smooth the reproduction of all of the
products of your research (simulations, data analyses, tables, figures, etc..). Similar features are offered by the Open Science Framework (https://osf.io/), Maarten recently participated in a workshop on research transparency organized by the Berkeley Institute for Transparency in the Social Sciences (www.bitss.org), all course material was accessible through OSF (https://osf.io/qh2nr/). Both platforms, OSF and Github nicely integrate and a key benefit is that they have been created to remain for a long time (unlike perhaps your university drive) <!-- ADD REFERENCE TO YOUR OSF PROJECT -->; (2) for one paper Jake simply included a Sweave document and data files into a compressed archive [@bowers2005dataverse] <!-- PERHAPS SAY ABIT MORE ABOUT WHAT SWEAVE IS IN PRINCIPLE-->; (3) for another more computing intensive paper Jake assembled a set of files that enabled reproduction of results using the
`make` system [@bowers2008dataverse] <!-- AGAIN A FEW WORDS ON WHAT PERHAPS LINK TO A WEBSITE? -->; (4) Jake also tried the "compendium"
approach [@gentleman2005reproducible,@gentleman2007statistical] which embeds an
academic paper within the R package system [@bowers2011dataverse]; The benefit of the compendium approach is that one is not required to have
access to a command line for `make`: the compendium is downloadable from within
R using `install.packages()` and is viewable using the `vignette()` function in
any operating system than runs R.^[Notice that Jake's reproduction archives and/or
instructions for using them are hosted on the
[Dataverse](http://thedata.org/book/learn-about-project), which is another
system designed to enhance academic collaboration across time and space.]  The
idea that one ought to be able to install and run and use an academic paper
just as one installs and uses statistical software packages is very attractive,
and we anticipate that it will become ever easier to turn papers into R packages
as creative and energetic folks turn their attention to the question of
reproducible research.^[See for example <http://r-pkgs.had.co.nz/>]
Finally, (5) Maarten has been using Dropbox for most of his collaborative projects (https://www.dropbox.com/). See comments above. As a tool for making files publicly available Dropbox is less useful. You can of course make replication files available through creating an public folder, but you will still need to refer to this folder on a particular project website (an example is Maarten's project on conflict and football https://www.dropbox.com/sh/wqosq99019mf04d/AAASTgJLkn6lmZJI_1veUK7ca?dl=0)

There are some noteworthy new developments also. For example, and similar to R Markdown/KnitR in R Studio, there is Markdoc for STATA users (https://github.com/haghish/MarkDoc). People at UC Berkeley have been developing a is a web app, Jupyter Notebook (http://jupyter.org/), that enables you to make and share documents that include text, code, equations and visualizations in one INCLUDE REF TO GITHUB - USING PYTHON. For researchers working across different platforms, there is Docker (https://www.docker.com/). There are also are great notes programs that help you communicate with collaborators. For example, Simplenote is very basic but allows you to share and change notes.

@@
